Flatiron cleaning

### 2021-10-08 Rmd/rdi analysis.Rmd
Fix bug removing line and clean interval of data to bind

  mutate(remove_linename_other = case_when(
    # keep these drugs in line 1 of upfront chemo and upfront surg
    linenumber_new == 1 & 
      drug_count == 1 &
       !str_detect(linename, "platin|taxel")                                 ~ "remove",
    linenumber_new == 1 &
      drug_count == 2 &
      (
      !(str_detect(linename, "platin") & str_detect(linename, "taxel") |
      str_detect(linename, "platin") & str_detect(linename, "Bev") |
      str_detect(linename, "taxel") & str_detect(linename, "Bev"))
      )                                                                      ~ "remove",
    linenumber_new == 1 &
      drug_count >= 3 &
      !(str_detect(linename, "platin") & str_detect(linename, "taxel") & str_detect(linename, "Bev"))

                                                                             ~ "remove",
    # weird one
    linenumber_new == 1 &
      drug_count >= 3 &
      str_detect(linename, "Niraparib|Gemcitabine")                          ~ "remove",

    # keep these drugs in line 2 of upfront chemo from when line switch
    linenumber_new == 2 &
      therapy == "Upfront Chemo" &
      drug_count == 1 &
       !str_detect(linename, "platin|taxel")  ~ "remove",

    linenumber_new == 2 &
      therapy == "Upfront Chemo" &
      drug_count == 2 &
      (
      !(str_detect(linename, "platin") & str_detect(linename, "taxel") |
      str_detect(linename, "platin") & str_detect(linename, "Bev") |
      str_detect(linename, "taxel") & str_detect(linename, "Bev"))
      ) ~ "remove",

    linenumber_new == 2 &
      therapy == "Upfront Chemo" &
      drug_count >= 3 &
      !(str_detect(linename, "platin") & str_detect(linename, "taxel") & str_detect(linename, "Bev"))
      # !str_detect(linename, "taxel") &
      # !str_detect(linename, "Bev")
                        ~ "remove",

@@

therapy_line %>% 
  filter(linenumber == 1#, treatment_sequence == "neo + surg + adj"
         ) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linename) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))

@@

```{r}
therapy_line %>% 
  filter(linenumber == 1#, treatment_sequence == "neo + surg + adj"
         ) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linename) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))

@@

### Number of patient who switch line because of surgery from line 1 neoadjuvant to an adjuvant line
```{r Number of patient who switch line}
therapy_line %>% 
  mutate(change_line = case_when(
    linenumber == 1 & 
      linenumber_new == 2 &
      chemotherapy_type == "Adjuvant" &
      treatment_sequence == "neo + surg + adj"     ~ 1,
    TRUE                                           ~ 0
  )) %>% 
  arrange(desc(change_line)) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(change_line) %>% 
  tbl_summary()
```

595 patients had part of their first line a switch of line but we have examples of patients for who the switched line last a really short time.  
Easy example (sorry for the poor looking table `r emo::ji("face_with_raised_eyebrow")`):  
In this example, 1 cycle is switched from the initial line 1 coded by flatiron to a linenumber_new 2

| linenumber (coded by flatiron) |  surgery |  linenumber_new that I coded   |  drug    |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | taxel    |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | platin   |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | taxel    |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | platin   |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | taxel    |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | platin   |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             1                  | taxel    |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  | happened |             2                  | platin   |  
| ------------------------------ | -------- | ------------------------------ | -------- |  
|             1                  |          |             2                  | taxel    |  
| ------------------------------ | -------- | ------------------------------ | -------- |

Here a real example of this type of patient : initial `linenumber` = 1 but had surgery in 09/2019 so part of this line in moved to `linenumber_new` = 2 and then coded as adjuvant line.  
So when we will calculate RDI for frontline adjuvant, it will only include this 1 cycle. <span style="color:red">**Is that correct?**</span>  
```{r Number of patient who switch line2, rows.print=30}
# Problem F0000AD999ABF but would be good for F003F2BEEDB37
therapy_line %>% 
  filter(str_detect(patientid, "F41E15E18E008")) %>% 
  arrange(patientid, episodedate) %>% 
  select(linenumber, linenumber_new, episodedate, chemotherapy_type, drugname)
```

Here an example a little more complicated because the patient had surgery 05/2017 and the same type of switch but then had a 3rd line (used to be line 2 in flatiron).    
So when we will calculate RDI for frontline adjuvant, it will only include this 1 cycle switched to adjuvant and not the longer line witch was coded 2 by flatiron. <span style="color:red">**Is that correct?**</span>  
```{r, rows.print=30}
therapy_line %>% 
  filter(str_detect(patientid, "FC9D28216FEBE")) %>% 
  arrange(patientid, episodedate) %>% 
  select(linenumber, linenumber_new, episodedate, chemotherapy_type, drugname)

# therapy_line %>% 
#   mutate(change_line = case_when(
#     linenumber == 1 & 
#       linenumber_new == 2 &
#       chemotherapy_type == "Adjuvant" &
#       treatment_sequence == "neo + surg + adj"     ~ 1,
#     TRUE                                           ~ 0
#   )) %>% 
#   filter(change_line == 1) %>% 
#   group_by(patientid) %>% 
#   mutate(firstdate = first(episodedate)) %>% 
#   mutate(lastdate = last(episodedate)) %>% 
#   ungroup() %>% 
#   mutate(time_of_switch_line = interval(start = firstdate, end = lastdate) / duration(n =1, units = "days")) %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   ggplot(aes(x=time_of_switch_line, fill = ..count..))+
#   geom_histogram(binwidth = 7)+
#   theme_minimal()+
#   ggforce::facet_zoom(xlim= c(0,60), zoom.size = 1) + 
#   scale_y_continuous(#breaks=c(0, 50, 100, 150),
#                      minor_breaks=c(10, 20, 30, 40)#,
#                      #labels = c(0, 50, 100, 150)
#   )
```
`r emo::ji("question_mark")`  
<span style="color:red">**So the bigger question is : Is it acceptable to include the patients who have a switched line?**</span>

@@

`r emo::ji("question_mark")`  
As you can see on the plots, the interval in days can be pretty big so :  
<span style="color:red">**Can we take a creatinine value to bind it to a drug when a year long as past between each other?  
Do we need to set up a maximun gap?  
Same question for height, weight, bsa and corresponding AUC/target_dose  
Let me know your advise, I will try to update that for our next meeting**</span> `r emo::ji("smiling_face")`

```{r target_auc and expected dose}
Treatment <- 
  # Merge with creatinine
  left_join(therapy_cycle, creatinine, by = "patientid") %>%
  mutate(interval_creat = abs(interval(start= creat_date, end= cycle_start_date)/                      
                          duration(n=1, unit="days"))) %>% 
  arrange(interval_creat) %>% 
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment, 
           cycle_start_date, .keep_all = TRUE) %>%

@@

  # Merge with weight
  left_join(. , weight, by = "patientid") %>%
  mutate(interval_weight = abs(interval(start= weight_date, end= cycle_start_date)/
                        duration(n=1, unit="days"))) %>%
  arrange(interval_weight) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
         cycle_start_date, .keep_all = TRUE) %>%

@@

  # Merge with body_surface_area
  left_join(., body_surface_area, by = "patientid") %>%
  mutate(interval_bsa = abs(interval(start= bsa_date, end= cycle_start_date)/
                          duration(n=1, unit="days"))) %>%
  arrange(interval_bsa) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
           cycle_start_date, .keep_all = TRUE) %>%
  # Calculate more bsa value with height and weight using Du Bois formula
  mutate(bsa_du_bois = 0.007184 * ((height*100)^0.725) * (weight^0.425)) %>% # get pretty close results 4269 patients result
  mutate(BSA = coalesce(BSA, bsa_du_bois)) %>%
  mutate(bsa_du_bois = 0.007184 * ((height_at_dx*100)^0.725) * (weight_at_dx^0.425)) %>% # get pretty close results 4269 patients result
  mutate(bsa_at_dx = coalesce(bsa_at_dx, bsa_du_bois)) %>%
  select(-bsa_du_bois) %>%

  # Merge with auc
  left_join(., areaUC, by = c("patientid", "drugname")) %>%
  mutate(interval_auc = abs(interval(start= auc_date, end= cycle_start_date)/ # use episode date to be more precise? No it's not better
                          duration(n=1, unit="days"))) %>% 
  arrange(interval_auc) %>% 
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment, 
           cycle_start_date, .keep_all = TRUE) %>%

@@

```{r plot interval}
Treatment %>% 
  ggplot(aes(x= interval_creat)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
# Treatment %>% 
#   ggplot(aes(x= interval_height)) +
#   geom_bar() +
#   coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_weight)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_bsa)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_auc)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
```

>>>>

  mutate(remove_linename_other = case_when(
    # keep these drugs in line 1 of upfront chemo and upfront surg
    # linenumber_new == 1 & 
    #   drug_count == 1 &
    #    !str_detect(linename, "platin|taxel")                                 ~ "remove",
    # linenumber_new == 1 &
    #   drug_count == 2 &
    #   (
    #   !(str_detect(linename, "platin") & str_detect(linename, "taxel") |
    #   str_detect(linename, "platin") & str_detect(linename, "Bev") |
    #   str_detect(linename, "taxel") & str_detect(linename, "Bev"))
    #   )                                                                      ~ "remove",
    # linenumber_new == 1 &
    #   drug_count >= 3 &
    #   !(str_detect(linename, "platin") & str_detect(linename, "taxel") & str_detect(linename, "Bev"))
    # 
    #                                                                          ~ "remove",
    # weird one
    linenumber_new == 1 &
      str_detect(linename, 
                 "parib|bine|Dox|T|F|Clinical|Cyclo|lumab|Tra|ximab|zole|Aza|trex|Gos|lide|Pem")                       
                                                                         ~ "remove",
    linenumber_new == 1 & linename == "Bevacizumab"                      ~ "remove",
    linenumber_new == 2 &
      therapy == "Upfront Chemo" &
      str_detect(linename, 
                 "parib|bine|Dox|T|F|Clinical|Cyclo")                       
                                                                         ~ "remove",
    linenumber_new == 2 &
      therapy == "Upfront Chemo" & linename == "Bevacizumab"             ~ "remove",
    # keep these drugs in line 2 of upfront chemo from when line switch
    # linenumber_new == 2 &
    #   therapy == "Upfront Chemo" &
    #   drug_count == 1 &
    #    !str_detect(linename, "platin|taxel")  ~ "remove",
    #   
    #   
    # linenumber_new == 2 &
    #   therapy == "Upfront Chemo" &
    #   drug_count == 2 &
    #   (
    #   !(str_detect(linename, "platin") & str_detect(linename, "taxel") |
    #   str_detect(linename, "platin") & str_detect(linename, "Bev") |
    #   str_detect(linename, "taxel") & str_detect(linename, "Bev"))
    #   ) ~ "remove",
    #   
    #   
    # linenumber_new == 2 &
    #   therapy == "Upfront Chemo" &
    #   drug_count >= 3 &
    #   !(str_detect(linename, "platin") & str_detect(linename, "taxel") & str_detect(linename, "Bev"))
    #   # !str_detect(linename, "taxel") &
    #   # !str_detect(linename, "Bev")
    #                     ~ "remove",

@@

# therapy_line %>% 
#   filter(linenumber == 1#, treatment_sequence == "neo + surg + adj"
#          ) %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(linename) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency"))
# 
# therapy_line %>% 
#   filter(linenumber_new == 2, therapy == "Upfront Chemo") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(linename) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency"))

@@

First table is frontline Neo, second table is frontline Adj.
```{r}
therapy_line %>% 
  filter(linenumber == 1#, treatment_sequence == "neo + surg + adj"
         ) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linename) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))

@@

### Number of patient who switch line because of surgery from line 1 neoadjuvant to an adjuvant line
```{r Number of patient who switch line}
therapy_line %>% 
  mutate(change_line = case_when(
    linenumber == 1 & 
      linenumber_new == 2 &
      chemotherapy_type == "Adjuvant" &
      therapy == "Upfront Chemo"                   ~ 1,
    TRUE                                           ~ 0
  )) %>% 
  arrange(desc(change_line)) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(change_line) %>% 
  tbl_summary()
```

Some patients had part of their first line switched to Adjuvant line. How long is this adjuvant line?
```{r}
therapy_line %>% 
  mutate(line_duration = interval(start = linestartdate, end = lineenddate)/ 
           duration(n = 1, units = "days")) %>% 
  filter(linenumber == 1 & linenumber_new == 2) %>% 
  select(patientid, linenumber, linenumber_new, linenumber_adj, 
         drugname, chemotherapy_type, therapy, episodedate, line_duration) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= line_duration))+
  geom_bar(color = "slateblue") +
  theme_minimal()+
  ggforce::facet_zoom(xlim = c(0, 200), zoom.size = 1)
```

<!-- Easy example (sorry for the poor looking table `r emo::ji("face_with_raised_eyebrow")`):   -->
<!-- In this example, 1 cycle is switched from the initial line 1 coded by flatiron to a linenumber_new 2 -->

<!-- <br> -->

<!-- | linenumber (coded by flatiron) |  surgery |  linenumber_new that I coded   |  drug    |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | taxel    |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | platin   |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | taxel    |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | platin   |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | taxel    |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | platin   |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             1                  | taxel    |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  | happened |             2                  | platin   |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->
<!-- |             1                  |          |             2                  | taxel    |   -->
<!-- | ------------------------------ | -------- | ------------------------------ | -------- |   -->

<!-- <br> -->

<!-- Here a real example of this type of patient : initial `linenumber` = 1 but had surgery in 09/2019 so part of this line in moved to `linenumber_new` = 2 and then coded as adjuvant line.   -->
<!-- So when we will calculate RDI for frontline adjuvant, it will only include this 1 cycle. <span style="color:red">**Is that correct?**</span>   -->
```{r Number of patient who switch line2, rows.print=30}
# Problem F0000AD999ABF but would be good for F003F2BEEDB37
# therapy_line %>% 
#   filter(str_detect(patientid, "F41E15E18E008")) %>% 
#   arrange(patientid, episodedate) %>% 
#   select(linenumber, linenumber_new, episodedate, chemotherapy_type, drugname)
```

<!-- Here an example a little more complicated because the patient had surgery 05/2017 and the same type of switch but then had a 3rd line (used to be line 2 in flatiron).     -->
<!-- So when we will calculate RDI for frontline adjuvant, it will only include this 1 cycle switched to adjuvant and not the longer line witch was coded 2 by flatiron. <span style="color:red">**Is that correct?**</span>   -->
```{r, rows.print=30}
# therapy_line %>% 
#   filter(str_detect(patientid, "FC9D28216FEBE")) %>% 
#   arrange(patientid, episodedate) %>% 
#   select(linenumber, linenumber_new, episodedate, chemotherapy_type, drugname)

# therapy_line %>% 
#   mutate(change_line = case_when(
#     linenumber == 1 & 
#       linenumber_new == 2 &
#       chemotherapy_type == "Adjuvant" &
#       treatment_sequence == "neo + surg + adj"     ~ 1,
#     TRUE                                           ~ 0
#   )) %>% 
#   filter(change_line == 1) %>% 
#   group_by(patientid) %>% 
#   mutate(firstdate = first(episodedate)) %>% 
#   mutate(lastdate = last(episodedate)) %>% 
#   ungroup() %>% 
#   mutate(time_of_switch_line = interval(start = firstdate, end = lastdate) / duration(n =1, units = "days")) %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   ggplot(aes(x=time_of_switch_line, fill = ..count..))+
#   geom_histogram(binwidth = 7)+
#   theme_minimal()+
#   ggforce::facet_zoom(xlim= c(0,60), zoom.size = 1) + 
#   scale_y_continuous(#breaks=c(0, 50, 100, 150),
#                      minor_breaks=c(10, 20, 30, 40)#,
#                      #labels = c(0, 50, 100, 150)
#   )
```
<!-- `r emo::ji("question_mark")`   -->
<!-- <span style="color:red">**So the bigger question is : Is it acceptable to include the patients who have a switched line?**</span> -->

@@

`r emo::ji("check")`  
We decided to take the creatine, weight, bsa measured within a month and in the same cycle of the administration.

```{r target_auc and expected dose}
Treatment <- 
  # Merge with creatinine
  left_join(therapy_cycle, creatinine, by = "patientid") %>%
  mutate(interval_creat = abs(interval(start= creat_date, end= cycle_start_date)/                      
                          duration(n=1, unit="days"))) %>% 
  mutate(interval_creat = case_when(
    interval_creat < 31 &
      creat_date > linestartdate &
      creat_date < lineenddate             ~ interval_creat,
    TRUE                                   ~ NA_real_
  )) %>% 
  arrange(interval_creat) %>% 
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment, 
           cycle_start_date, .keep_all = TRUE) %>%

@@

  # Merge with weight
  left_join(. , weight, by = "patientid") %>%
  mutate(interval_weight = abs(interval(start= weight_date, end= cycle_start_date)/
                        duration(n=1, unit="days"))) %>%
  mutate(interval_weight = case_when(
    interval_weight < 31 &
      weight_date > linestartdate &
      weight_date < lineenddate             ~ interval_weight,
    TRUE                                    ~ NA_real_
  )) %>% 
  arrange(interval_weight) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
         cycle_start_date, .keep_all = TRUE) %>%

@@

  # Merge with body_surface_area
  left_join(., body_surface_area, by = "patientid") %>%
  mutate(interval_bsa = abs(interval(start= bsa_date, end= cycle_start_date)/
                          duration(n=1, unit="days"))) %>%
  mutate(interval_bsa = case_when(
    interval_bsa < 31 &
      bsa_date > linestartdate &
      bsa_date < lineenddate               ~ interval_bsa,
    TRUE                                   ~ NA_real_
  )) %>% 
  arrange(interval_bsa) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
           cycle_start_date, .keep_all = TRUE) %>%
  # Calculate more bsa value with height and weight using Du Bois formula
  mutate(bsa_du_bois = 0.007184 * ((height*100)^0.725) * (weight^0.425)) %>% # get pretty close results 4269 patients result
  mutate(BSA = coalesce(BSA, bsa_du_bois)) %>%
  mutate(bsa_du_bois = 0.007184 * ((height_at_dx*100)^0.725) * (weight_at_dx^0.425)) %>% # get pretty close results 4269 patients result
  mutate(bsa_at_dx = coalesce(bsa_at_dx, bsa_du_bois)) %>%
  select(-bsa_du_bois) %>%

  # Merge with auc
  left_join(., areaUC, by = c("patientid", "drugname")) %>%
  mutate(interval_auc = abs(interval(start= auc_date, end= cycle_start_date)/ # use episode date to be more precise? No it's not better
                          duration(n=1, unit="days"))) %>% 
  mutate(interval_auc = case_when(
    interval_auc < 31 &
      auc_date > linestartdate &
      auc_date < lineenddate               ~ interval_auc,
    TRUE                                   ~ NA_real_
  )) %>% 
  arrange(interval_auc) %>% 
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment, 
           cycle_start_date, .keep_all = TRUE) %>%

@@

```{r plot interval}
Treatment %>% 
  ggplot(aes(x= interval_creat)) +
  geom_bar() #+
  # coord_cartesian(ylim=c(0,50))
# Treatment %>% 
#   ggplot(aes(x= interval_height)) +
#   geom_bar() +
#   coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_weight)) +
  geom_bar() #+
  # coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_bsa)) +
  geom_bar() #+
  # coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_auc)) +
  geom_bar() #+
  # coord_cartesian(ylim=c(0,50))
```
